---
title: "SLP Estimations: Data Analysis"
output: html_notebook
---

This is the analysis script for the manuscript entitled "The reliability and validity of SLPs' esitmations of speech intelligibility in dysarthria". Much of the analysis and code is the same from the 2021 ASHA Convention poster presentation. However, some of the data visualizations and tables have changed.

# Packages

```{r}

library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
library(corrr) # install.packages("corrr")
library(performance) # install.packages("performance")
library(furniture) #install.packages("furniture")
library(see) # install.packages("see")
library(patchwork) # install.packages("patchwork")
library(ggridges) #install.packages("ggridges")
library(rmcorr) #install.packages("rmcorr")
library(irr) #install.packages("irr")
library(ggpubr) #install.packages("ggpubr")
library(RColorBrewer) #install.packages("RColorBrewer")
library(extrafont) # install.packages("extrafont")
library(gt) #install.packages("gt")
#library(glue) #install.packages("glue")

```

# Importing the datasets

This chunk contains all of the dataframes needed for the stat analysis.
```{r}
# This data is used for question #1
IntSpeakerRatings <- rio::import("Shared Data/SpeakerIntelligibility.csv")

# This data is used for question #2 & #4
slp_Demographics <- rio::import("Shared Data/SLP_Demographics.csv")
slp_Correlation <- rio::import("Shared Data/SLP_Correlation.csv")

# This data is used for question #3
naive_Demographics <- rio::import("Shared Data/NaiveListener_Demographics.csv")
naive_Correlation <- rio::import("Shared Data/NaiveListener_Correlation.csv")

```

# Descriptive Stats

## SLP Demographics

```{r}

slp_gender <- slp_Demographics %>%
  dplyr::select(gender) %>%
  dplyr::group_by(gender) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "gender") %>%
  dplyr::mutate(Group = "Gender")

slp_race <- slp_Demographics %>%
  dplyr::select(race) %>%
  dplyr::group_by(race) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "race") %>%
  dplyr::mutate(Group = "Race")

slp_ethnicity <- slp_Demographics %>%
  dplyr::select(ethnicity) %>%
  dplyr::group_by(ethnicity) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "ethnicity") %>%
  dplyr::mutate(Group = "Ethnicity")

slp_demo1 <- dplyr::full_join(slp_gender, slp_race)

slp_demo1 <- dplyr::full_join(slp_demo1, slp_ethnicity)
  
slp_demo_table <- slp_demo1 %>%
  gt::gt(rowname_col = "Category", groupname_col = "Group")

slp_demo2 <- slp_Demographics %>%
  dplyr::select(age, yearsExp) %>%
  purrr::map_df(.f = ~ broom::tidy(summary(.x)), .id = "Variable") %>%
  dplyr::select(!c(q1, median, q3))

  

```

## Naive Listener Demographics

```{r}

naive_Demographics %>%
  furniture::table1(age, gender, race, ethnicity, famSLHDs)

```

## SLP descriptives

```{r}

# Print mean and SD
MeanSLPInt <- slp_Correlation %>%
  dplyr::group_by(Speaker) %>%
  dplyr::summarize(VAS = mean(slp_VAS, na.rm = T), VAS_SD = sd(slp_VAS, na.rm = T),
                   EST = mean(slp_EST, na.rm =T ), EST_SD = sd(slp_EST, na.rm = T))

MeanSLPInt

```

## Naive Listener Descriptives

```{r}

naive_Correlation %>%
  dplyr::group_by(SpeakerID) %>%
  dplyr::summarize(OT = mean(naive_OT), OT_SD = sd(naive_OT), VAS = mean(naive_VAS), VAS_SD = sd(naive_VAS))

```


## Descriptives Plot

```{r}

intRatings_Long1 <- IntSpeakerRatings %>%
  dplyr::select(SpeakerID, slp_VAS, slp_EST) %>%
  pivot_longer(cols = slp_EST:slp_VAS,
               names_to = "RatingType",
               values_to = "Ratings")

intRatings_Long1 <- slp_Correlation %>%
  dplyr::rename(SpeakerID = Speaker) %>%
  dplyr::select(SpeakerID, slp_VAS, slp_EST) %>%
  pivot_longer(cols = slp_EST:slp_VAS,
               names_to = "RatingType",
               values_to = "Ratings")

intRatings_Long2 <- naive_Correlation %>%
  dplyr::select(SpeakerID, naive_VAS, naive_OT) %>%
    pivot_longer(cols = naive_VAS:naive_OT,
               names_to = "RatingType",
               values_to = "Ratings")

intRatings_Long <- base::rbind(intRatings_Long1, intRatings_Long2) %>%
  dplyr::mutate(RatingType = as.factor(RatingType)) %>%
  dplyr::mutate(RatingType = recode_factor(RatingType,
                                           "naive_VAS" = "Naive (VAS)",
                                           "naive_OT" = "Naive (Transcriptions)",
                                           "slp_VAS" = "SLP (VAS)",
                                           "slp_EST" = "SLP (Estimations)")) %>%
  dplyr::mutate(RatingType = fct_relevel(RatingType,
                                                "Naive (Transcriptions)",
                                                "Naive (VAS)",
                                                "slp_VAS" = "SLP (VAS)",
                                                "slp_EST" = "SLP (Estimations)"))

## custom colors
my_pal <- rcartocolor::carto_pal(n = 7, name = "Teal")[c(7, 5, 3, 2, 4)]

## Theme ----
theme_set(theme_void(base_family = "Roboto"))

theme_update(
  axis.text.x = element_text(color = "black", face = "bold", size = 26, 
                             margin = margin(t = 6)),
  axis.text.y = element_text(color = "black", size = 22, hjust = 1, 
                             margin = margin(r = 6), family = "Roboto Mono"),
  axis.line.x = element_line(color = "black", size = 1),
  panel.grid.major.y = element_line(color = "grey90", size = .6),
  plot.background = element_rect(fill = "white", color = "white"),
  plot.margin = margin(rep(20, 4))
)


# Difference Plots: Word Trans ----                                                       
ggplot(data = intRatings_Long,
       aes(x = RatingType,
           y = Ratings,
           color = RatingType,
           fill = RatingType)) + 
  ggdist::stat_halfeye(
    adjust = 1.3, 
    width = .6, 
    .width = 0, 
    justification = -.2,
    point_colour = NA,
    outline_bars = T
  ) + 
  scale_alpha_discrete(range = c(.4,.9), guide = "none") +
  geom_boxplot(
    width = .15, 
    outlier.shape = NA,
    fill = "white"
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .3
  ) +
  coord_cartesian(xlim = c(0, 100), clip = "off") +
  coord_flip() +
  scale_color_manual(values = my_pal, guide = "none") +
  scale_fill_manual(values = my_pal, guide = "none") +
  labs(title = "Intelligibility Ratings") +
  ylab("Intelligibility") +
  xlab("") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size = 13, 
        family="Arial Bold"))

# ggsave("Plots/IntRatingsDistribution.png", plot = last_plot(),height = 5, width = 7, units = "in", scale = .8)

```


# Research Question #1
Q1: Are SLP estimates of speech intelligibility predictive of naive listener estimations?

Since the speakers are the common variable between the SLP and Naive listener task, intelligibility scores for each method were calculated by speaker.

## Correlation Matrix

This first block explores the correlation matrix of SLP and Naive listener intelligibility ratings. 

```{r}

IntSpeakerRatings %>%
  dplyr::select(!SpeakerID) %>%
  corrr::correlate(method = "pearson", use = "pairwise.complete.obs", quiet = T)

```

## Modeling naive OT scores 

This next block is modeling SLP estimation as predictor of naive listener OT.

```{r}

SLP_ESTmodel <- lm(naive_OT ~ slp_EST, data= IntSpeakerRatings)
summary(SLP_ESTmodel)
performance::check_model(SLP_ESTmodel)

```


This next block is examining SLP VAS as a predictor of naive listener OT.

```{r}

SLP_VASmodel <- lm(naive_OT ~ slp_VAS, data = IntSpeakerRatings)
summary(SLP_VASmodel)
performance::check_model(SLP_VASmodel)

```


Naive OT Model comparisons

```{r}
performance::compare_performance(SLP_ESTmodel, SLP_VASmodel, rank = T)

performance::test_performance(SLP_ESTmodel, SLP_VASmodel)
```


## Additional Analysis: Modeling naive VAS scores

This next block is examining SLP EST as a predictor of naive listener VAS.

```{r}
SLP_ESTmodel_VAS <- lm(naive_VAS ~ slp_EST, data = IntSpeakerRatings)
summary(SLP_ESTmodel_VAS)
performance::check_model(SLP_ESTmodel_VAS)
```

This block examines SLP VAS as a predictor of naive listener VAS.

```{r}
SLP_VASmodel_VAS <- lm(naive_VAS ~ slp_VAS, data = IntSpeakerRatings)
summary(SLP_VASmodel_VAS)
performance::check_model(SLP_VASmodel_VAS)
```

Naive VAS model comparison

```{r}
performance::compare_performance(SLP_ESTmodel_VAS, SLP_VASmodel_VAS, rank = T)

performance::test_performance(SLP_ESTmodel_VAS, SLP_VASmodel_VAS)
```

## RQ1 Plot

```{r}

# custom colors
my_pal <- c("#3C1642", "#086375", "#1DD3B0", "#AFFC41")

IntSpeakerRatings_Long <- IntSpeakerRatings %>%
  tidyr::pivot_longer(cols = slp_EST:slp_VAS, names_to = "Method",values_to = "rating") %>%
  dplyr::mutate(Method = recode(Method, slp_EST = "Estimation", slp_VAS = "VAS"))

NaiveOT_plot <- IntSpeakerRatings_Long %>%
  ggplot() +
  aes(x = rating, 
      y = naive_OT,
      color = Method,
      ymin = 0,
      ymax = 100,
      xmin = 0,
      xmax = 100) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  geom_abline(intercept = 0, slope = 1, linetype = 2) +
  ylab("Naïve Intelligibilty (Transcription)") +
  labs(color = "SLP Rating Method") +
  scale_color_manual(values = my_pal) +
  coord_equal() +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.title.x=element_blank())

ggsave("Plots/RQ1_ScatterPlot.png", plot = last_plot(),height = 5, width = 5, units = "in")

```


# Research Question #2
Q2: Is there a strong relationship between SLP intelligibility approximations with SLP VAS intelligibility ratings?

```{r}

rmcorr::rmcorr(slpID, slp_VAS, slp_EST, slp_Correlation)

```


# Research Question #3
Q3: Is there a strong relationship between naive listener OT scores with their VAS intelligibility estimation?

```{r}

rmcorr::rmcorr(ListenerID, naive_OT, naive_VAS, naive_Correlation)

```

# Research Question #4
Q4: How reliable are these intelligibility measures (VAS ratings, approximations, and OT scores), as determined by intrarater and interrater agreement?

## Intrarater Reliability

### SLP VAS Ratings

```{r}

cor.test(slp_Correlation$slp_VAS, slp_Correlation$slp_VAS.Rel, method = "pearson", use = "complete.obs")

```

### SLP EST Ratings

```{r}

cor.test(slp_Correlation$slp_EST, slp_Correlation$slp_EST.Rel, method = "pearson", use = "complete.obs")

```

## Interrater Reliability 

### SLPs VAS rating

```{r}

slpVAS_Inter <- slp_Correlation %>%
  dplyr::select(!c(slp_VAS.Rel, slp_EST, slp_EST.Rel))

slpVAS_Inter_wide <- tidyr::pivot_wider(slpVAS_Inter, names_from = slpID, values_from = slp_VAS)

slpVAS_Inter_wide <- slpVAS_Inter_wide %>%
  dplyr::select(!Speaker)

k <- 1
while(k <= ncol(slpVAS_Inter_wide)){
  slpVAS_Inter_wide[is.na(slpVAS_Inter_wide[,k]), k] <- colMeans(slpVAS_Inter_wide[,k], na.rm = TRUE)
  k <- k + 1
}

irr::icc(slpVAS_Inter_wide, model = "twoway",
  type="agreement",
  unit="average", r0=0, conf.level=0.95)

irr::icc(slpVAS_Inter_wide, model = "twoway",
  type="agreement",
  unit="single", r0=0, conf.level=0.95)
```


### SLPs Estimations

```{r}

slpEST_Inter <- slp_Correlation %>%
  dplyr::select(!c(slp_EST.Rel, slp_VAS, slp_VAS.Rel))

slpEST_Inter_wide <- tidyr::pivot_wider(slpEST_Inter, names_from = slpID, values_from = slp_EST)

slpEST_Inter_wide <- slpEST_Inter_wide %>%
  dplyr::select(!Speaker)

irr::icc(slpEST_Inter_wide, model = "twoway",
  type="agreement",
  unit="average", r0=0, conf.level=0.95)

irr::icc(slpEST_Inter_wide, model = "twoway",
  type="agreement",
  unit="single", r0=0, conf.level=0.95)

```

# Extra Data Visualization

## Creating Mean SLP Int plot

```{r}

MeanInt_plot <- MeanSLPInt %>%
  ggplot() +
  aes(x = Method, 
      y = Intelligibility,
      group = Speaker,
      color = Speaker) +
  geom_point() +
  geom_line() +
  theme_classic()

MeanInt_plot

ggsave(filename = "Plots/SpeakerInt_Plot.png", plot = MeanInt_plot, height = 10, width = 10, unit = "in")

```


## SLP Distribution Plot 

```{r}

# SLP Distribution Plot for Estimates

distributionData <- slp_Correlation %>%
  dplyr::mutate(Speaker = fct_relevel(Speaker, levels = "AF9", "ALSF6", "HDM8", "ALSF9", "HDM10", "ALSF7", "AM1", "AM5", "AM3", "PDM12", "PDM9", "HDF5", "ALSM1", "HDF1", "ALSM4", "AF1", "PDM10", "PDF5", "PDF7", "HDM11"))

slp_EST.DistPlot <- distributionData %>%
  ggplot() +
  aes(x = slp_EST, 
      y = Speaker,
      fill = Speaker,
      color = Speaker) +
  xlim(c(0,100)) +
  xlab("Estimation Rating") +
  geom_density_ridges(alpha = 0.5) +
  theme_classic() +
  theme(legend.position = "none")

# SLP Distribution Plot for VAS

distributionData1 <- slp_Correlation %>%
  dplyr::mutate(Speaker = fct_relevel(Speaker, levels = "AF9", "HDM10", "ALSF6", "HDM8", "ALSF9", "AM1", "AM3", "ALSF7", "AM5", "PDM12", "HDF5", "ALSM1", "ALSM4", "HDF1", "AF1", "PDF7", "PDM9", "PDM10", "PDF5", "HDM11"))

slp_VAS.DistPlot <- distributionData1 %>%
  ggplot() +
  aes(x = slp_VAS, 
      y = Speaker,
      fill = Speaker,
      color = Speaker) +
  xlim(c(0,100)) +
  xlab("Visual Analog Scale Rating") +
  geom_density_ridges(alpha = 0.5) +
  theme_classic() +
  theme(legend.position = "none")

# Creating Final plot

DistributionPlot <- ggpubr::ggarrange(slp_EST.DistPlot, slp_VAS.DistPlot, ncol = 2, labels = c("A", "B"), nrow = 1) 

DistributionPlot <- annotate_figure(DistributionPlot, top = text_grob("SLP Intelligibility Ratings per Speaker", 
               face = "bold", size = 14))

ggsave(filename = "Plots/Distribution_Plot.png", plot = DistributionPlot, height = 5, width = 10, unit = "in")

# removing plots from environment

rm(distributionData, distributionData1, slp_EST.DistPlot, slp_VAS.DistPlot)

```

