---
title: "SLP Estimations: Data Analysis"
output: html_notebook
---

This is the analysis script for the manuscript entitled "The reliability and validity of SLPs' esitmations of speech intelligibility in dysarthria". Much of the analysis and code is the same from the 2021 ASHA Convention poster presentation. However, some of the data visualizations and tables have changed.

# Packages

```{r}

library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
library(performance) # install.packages("performance")
library(see) # install.packages("see")
library(patchwork) # install.packages("patchwork")
library(rmcorr) #install.packages("rmcorr")
library(irr) #install.packages("irr")
library(Hmisc) #install.packages("Hmisc")
library(ggpubr) #install.packages("ggpubr")
library(extrafont) # install.packages("extrafont")
library(gt) #install.packages("gt")
library(sjPlot) #install.packages("sjPlot")

```

# Importing the datasets

This chunk contains all of the dataframes needed for the statistical analysis.

```{r}

# This data is used for question #1 & 2
slp_Demographics <- rio::import("Shared Data/SLP_Demographics.csv")
slp_Correlation <- rio::import("Shared Data/SLP_Correlation.csv")

# This data is used for question #3
IntSpeakerRatings <- rio::import("Shared Data/SpeakerIntelligibility.csv")

# Naïve Listener Data
naive_Demographics <- rio::import("Shared Data/NaiveListener_Demographics.csv")
naive_Correlation <- rio::import("Shared Data/NaiveListener_Correlation.csv")

```

# Descriptive Statistics

## SLP Demographics

```{r}
 
#creating separate data frames for SLP gender, race, and ethnicity demographic information

slp_gender <- slp_Demographics %>%
  dplyr::select(gender) %>%
  dplyr::group_by(gender) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "gender") %>%
  dplyr::mutate(Group = "Gender",
                Category = fct_relevel(Category, c("Woman", "Man", "Prefer not to say"))) %>%
  dplyr::arrange(Category)

slp_race <- slp_Demographics %>%
  dplyr::select(race) %>%
  dplyr::group_by(race) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "race") %>%
  dplyr::mutate(Group = "Race")

slp_ethnicity <- slp_Demographics %>%
  dplyr::select(ethnicity) %>%
  dplyr::group_by(ethnicity) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "ethnicity") %>%
  dplyr::mutate(Group = "Ethnicity")

#merging these separate data frames together

slp_demo1 <- dplyr::full_join(slp_gender, slp_race)

slp_demo1 <- dplyr::full_join(slp_demo1, slp_ethnicity)

#creating the SLP demographic tables 

## Race, gender, and ethicity information

slp_demo_table <- slp_demo1 %>%
  gt::gt(rowname_col = "Category", groupname_col = "Group")

## SLP Age and Years experience information

slp_demo2 <- slp_Demographics %>%
  dplyr::select(c(age, yearsExp)) %>%
  tidyr::pivot_longer(cols = c(age, yearsExp), names_to = "Variable", values_to = "Value") %>%
  dplyr::mutate(Variable = recode(Variable, age = "Age", yearsExp = "Years Experience")) %>%
  dplyr::group_by(Variable) %>%
  dplyr::summarize(Mean = mean(Value), SD = sd(Value), Min = min(Value), Max = max(Value)) %>%
  gt::gt() %>%
  gt::fmt_number(columns = c(Mean, SD), decimals = 2)

# removing extra items from the environment that are no longer needed

rm(slp_gender, slp_race, slp_ethnicity, slp_demo1)

```

## Naive Listener Demographics

```{r}

#creating separate data frames for naive listener gender, race, ethnicity, and familiarity with speech language and hearing disorders (famSLHDs) information

naive_gender <- naive_Demographics %>%
  dplyr::select(gender) %>%
  dplyr::group_by(gender) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "gender") %>%
  dplyr::mutate(Group = "Gender",
                Category = fct_relevel(Category, c("Woman", "Man", "Prefer not to say"))) %>%
  dplyr::arrange(Category)

naive_race <- naive_Demographics %>%
  dplyr::select(race) %>%
  dplyr::group_by(race) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "race") %>%
  dplyr::mutate(Group = "Race")

naive_ethnicity <- naive_Demographics %>%
  dplyr::select(ethnicity) %>%
  dplyr::group_by(ethnicity) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "ethnicity") %>%
  dplyr::mutate(Group = "Ethnicity")

naive_famSLHDs <- naive_Demographics %>%
  dplyr::select(famSLHDs) %>%
  dplyr::group_by(famSLHDs) %>%
  dplyr::summarize(Frequency = n()) %>%
  dplyr::rename(Category = "famSLHDs") %>%
  dplyr::mutate(Group = "Familiar with Comm Disorders")

# Merging these separate data frames together

naive_demo1 <- dplyr::full_join(naive_gender, naive_race)

naive_demo2 <- dplyr::full_join(naive_demo1, naive_ethnicity)

naive_demo3 <- dplyr::full_join(naive_demo2, naive_famSLHDs)

#creating demographic table

naive_demo_table <- naive_demo3 %>%
  gt::gt(rowname_col = "Category", groupname_col = "Group")

#calculating demographic age information

naive_age <- naive_Demographics %>%
  dplyr::summarize(Age_mean = mean(age), Age_sd = sd(age), Age_min = min(age), Age_max = max(age))

# removing extra items from the environment that are no longer needed

rm(naive_gender, naive_race, naive_ethnicity, naive_famSLHDs, slp_demo1, naive_demo2, naive_demo3)

```

## SLP and Naive Listener Descriptive Measures by Speaker (including SD here)

```{r}

# Two data frames to calculate descriptive statistics for SLP and naive listener intelligibility measures per speaker
## Note: The InSpeakerRatings data frame already has the mean intelligibility measure values for all speakers (SLP VAS, estimations, and naive OT). However, standard deviation and range is not included. So it was recalculated here. 

SLP_data <- slp_Correlation %>%
  dplyr::select(!c(slp_VAS.Rel, slp_EST.Rel)) %>%
  tidyr::pivot_longer(cols = slp_VAS:slp_EST, names_to = "Method", values_to = "Int") %>%
  dplyr::mutate(Method = ifelse(Method == "slp_EST", "SLP Estimation", "SLP VAS Rating")) %>%
  dplyr::group_by(Speaker, Method) %>%
  dplyr::summarize(Mean = mean(Int, na.rm = T), SD = sd(Int, na.rm = T),
                   Min = min(Int, na.rm =T ), Max = max(Int, na.rm = T))

naive_data <- naive_Correlation %>%
  dplyr::select(!naive_VAS) %>%
  dplyr::group_by(SpeakerID) %>%
  dplyr::summarize(Mean = mean(naive_OT), SD = sd(naive_OT), Min = min(naive_OT), Max = max(naive_OT)) %>%
  dplyr::mutate(Method = "Naïve OT") %>%
  dplyr::rename(Speaker = "SpeakerID")

# Merging the two data frames

Descriptives_table <- full_join(SLP_data, naive_data)

# removing extra items from the environment

rm(SLP_data, naive_data)

```


## Descriptives Plot

```{r}

# Creating long-form data frame for descriptives plot

intRatings_Long1 <- slp_Correlation %>%
  dplyr::rename(SpeakerID = Speaker) %>%
  dplyr::select(c(SpeakerID, slp_VAS, slp_EST)) %>%
  pivot_longer(cols = slp_EST:slp_VAS,
               names_to = "RatingType",
               values_to = "Ratings")

intRatings_Long2 <- naive_Correlation %>%
  dplyr::select(c(SpeakerID, naive_OT)) %>%
  pivot_longer(cols = naive_OT,
               names_to = "RatingType",
               values_to = "Ratings")

plot_data <- base::rbind(intRatings_Long1, intRatings_Long2) %>%
  dplyr::mutate(RatingType = as.factor(RatingType)) %>%
  dplyr::mutate(RatingType = recode_factor(RatingType,
                                           "naive_OT" = "Naive (Transcriptions)",
                                           "slp_VAS" = "SLP (VAS)",
                                           "slp_EST" = "SLP (Estimations)")) %>%
  dplyr::mutate(RatingType = fct_relevel(RatingType,
                                                "Naive (Transcriptions)",
                                                "slp_VAS" = "SLP (VAS)",
                                                "slp_EST" = "SLP (Estimations)"))

# custom colors (used for all plots in this script)

my_pal <- c("#086375", "#B147C2", "#3C1642", "#0FB1D2")

# Difference Plots 

plot_data %>%
  ggplot() +
       aes(x = RatingType,
           y = Ratings,
           color = RatingType,
           fill = RatingType) +
  ## adding distribution plots 
  ggdist::stat_halfeye(
    adjust = 1.3, 
    width = .6, 
    .width = 0, 
    justification = -.2,
    point_colour = NA,
    outline_bars = T
  ) + 
  scale_alpha_discrete(range = c(.4,.9), guide = "none") +
  ## Adding boxplot under distributions
  geom_boxplot(
    width = .15, 
    outlier.shape = NA,
    fill = "white"
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .3
  ) +
  coord_cartesian(xlim = c(0, 100), clip = "off") +
  coord_flip() +
  scale_color_manual(values = my_pal, guide = "none") +
  scale_fill_manual(values = my_pal, guide = "none") +
  ylab("Intelligibility") +
  xlab("") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(text=element_text(size = 13, 
        family="Arial Bold"))

# Saving Distribution plot

ggsave("Plots/Distribution_plot.png", plot = last_plot(),height = 5, width = 7, units = "in", scale = .8)

# removing extra items from the environment

rm(intRatings_Long1, intRatings_Long2)

```


# Research Question #1

Q1: How reliable are these intelligibility measures (VAS ratings, approximations, and OT scores), as determined by intrarater and interrater agreement?

## Intrarater Reliability

### SLP VAS Ratings

```{r}

cor.test(slp_Correlation$slp_VAS, slp_Correlation$slp_VAS.Rel, method = "pearson", use = "complete.obs")

```

### SLP EST Ratings

```{r}

cor.test(slp_Correlation$slp_EST, slp_Correlation$slp_EST.Rel, method = "pearson", use = "complete.obs")

```

## Interrater Reliability 

### SLPs VAS rating

```{r}

slpVAS_Inter <- slp_Correlation %>%
  dplyr::select(!c(slp_VAS.Rel, slp_EST, slp_EST.Rel))

slpVAS_Inter_wide <- tidyr::pivot_wider(slpVAS_Inter, names_from = slpID, values_from = slp_VAS)

slpVAS_Inter_wide <- slpVAS_Inter_wide %>%
  dplyr::select(!Speaker)

k <- 1
while(k <= ncol(slpVAS_Inter_wide)){
  slpVAS_Inter_wide[is.na(slpVAS_Inter_wide[,k]), k] <- colMeans(slpVAS_Inter_wide[,k], na.rm = TRUE)
  k <- k + 1
}

irr::icc(slpVAS_Inter_wide, model = "twoway",
  type="agreement",
  unit="average", r0=0, conf.level=0.95)

irr::icc(slpVAS_Inter_wide, model = "twoway",
  type="agreement",
  unit="single", r0=0, conf.level=0.95)

# removing extra items from the environment

rm(slpVAS_Inter)

```

### SLPs Estimations

```{r}

slpEST_Inter <- slp_Correlation %>%
  dplyr::select(!c(slp_EST.Rel, slp_VAS, slp_VAS.Rel))

slpEST_Inter_wide <- tidyr::pivot_wider(slpEST_Inter, names_from = slpID, values_from = slp_EST)

slpEST_Inter_wide <- slpEST_Inter_wide %>%
  dplyr::select(!Speaker)

irr::icc(slpEST_Inter_wide, model = "twoway",
  type="agreement",
  unit="average", r0=0, conf.level=0.95)

irr::icc(slpEST_Inter_wide, model = "twoway",
  type="agreement",
  unit="single", r0=0, conf.level=0.95)

# removing extra items from the environment

rm(slpEST_Inter)

```

# Research Question #2

## Repeated Measures Correlation Analysis

Q2: Is there a strong relationship between SLP intelligibility percent estimates with SLP VAS intelligibility ratings?

```{r}

rmcorr::rmcorr(slpID, slp_VAS, slp_EST, slp_Correlation)

```

## Scatterplot

```{r}

slp_Correlation %>%
  ggplot() +
  aes(x = slp_VAS,
      y = slp_EST) +
  geom_abline(intercept = 0, slope = 1, linetype = 2) +
  geom_line(method = "lm", stat ="smooth", color = my_pal[2], aes(group = slpID), alpha = .4) +
  geom_smooth(method = "lm", se = F, color = my_pal[2]) +
  xlab("SLP VAS Rating") +
  ylab("SLP Percent Estimates") +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  theme_classic() +
  theme(aspect.ratio = 1) 

ggsave("Plots/RQ2_ScatterPlot.png", plot = last_plot(), height = 5, width = 5, units = "in", scale = .8)

```


# Research Question #3

Q1: Are SLP estimates of speech intelligibility predictive of naive listener estimations?

Since the speakers are the common variable between the SLP and Naive listener task, the mean SLP VAS rating, SLP estimation, and naive listener OT scores were calculated for each of the 20 speakers.

## Correlation Matrix

This first block explores the correlation matrix of SLP and Naive listener intelligibility ratings. 

```{r}

# Creating a correlation Matrix

CorrMatrix <- IntSpeakerRatings %>%
  dplyr::select(!SpeakerID) %>%
  as.matrix() %>%
  Hmisc::rcorr() 

# Testing to see if all coefficients are significant at the .05 level

CorrMatrixP <- CorrMatrix$P < .05

CorrMatrix <- CorrMatrix$r

# Creating a correlation matrix table to export and save

CorrMatrix <- CorrMatrix %>%
  as.data.frame() %>%
  dplyr::mutate(naive_VAS = as.numeric(naive_VAS),
                naive_OT = as.numeric(naive_OT),
                slp_EST = as.numeric(slp_EST),
                slp_VAS = as.numeric(slp_VAS)) %>%
  gt::gt() %>%
  gt::fmt_number(columns = c(naive_VAS, naive_OT, slp_EST, slp_VAS), decimals = 2) %>%
  gt::gtsave("Tables/CorrMatrix.html")

```

## Modeling naive OT scores 

### Modeling SLP estimation as predictor of naive listener OT.

```{r}

SLP_ESTmodel <- lm(naive_OT ~ slp_EST, data= IntSpeakerRatings)
summary(SLP_ESTmodel)
performance::check_model(SLP_ESTmodel)
sjPlot::tab_model(SLP_ESTmodel, file = "Tables/Estimation_model.html")


```


### Modeling SLP VAS as a predictor of naive listener OT.

```{r}

SLP_VASmodel <- lm(naive_OT ~ slp_VAS, data = IntSpeakerRatings)
summary(SLP_VASmodel)
performance::check_model(SLP_VASmodel)
sjPlot::tab_model(SLP_VASmodel, file = "Tables/VAS_model.html")

```


## RQ3 Plot

```{r}

# creating long form data frame for plot

IntSpeakerRatings_Long <- IntSpeakerRatings %>%
  tidyr::pivot_longer(cols = slp_EST:slp_VAS, names_to = "Method",values_to = "rating") %>%
  dplyr::mutate(Method = recode(Method, slp_EST = "Estimation", slp_VAS = "VAS")) %>%
  dplyr::mutate(Method = as.factor(Method)) %>%
  dplyr::mutate(Method = fct_relevel(Method, "VAS", "Estimation"))

# Creating RQ3 plot

IntSpeakerRatings_Long %>%
  ggplot() +
  aes(x = rating, 
      y = naive_OT,
      color = Method,
      shape = Method) +
  #adding data pairs and fitted regression line
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  #adding reference line to indicate theoretical perfict fit
  geom_abline(intercept = 0, slope = 1, linetype = 2) +
  ylab("Naïve Listener Intelligibilty (OT Score)") +
  labs(color = "SLP Intelligibility Measure", shape = "SLP Intelligibility Measure") +
  scale_color_manual(values = my_pal[2:3]) +
  scale_shape_manual(values = c(1, 16)) +
  #equalize the x and y axes
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.title.x=element_blank(), 
        aspect.ratio = 1)

#Saving the scatterplot
ggsave("Plots/RQ3_ScatterPlot.png", plot = last_plot(), height = 5, width = 5, units = "in", scale = .8)

```


# Additional Analysis

## Relationship between Naïve OT scores and Naïve VAS Ratings of Intelligibility

### Repeated Measures Correlation

To test whether our results are consistent with previous research, we investigated the relationship between naïve OT scores and VAS ratings of intelligibility.

```{r}

rmcorr::rmcorr(ListenerID, naive_OT, naive_VAS, naive_Correlation)

```

### Scatterplot

To depict relationship between the two naïve listener intelligibility measures

```{r}

naive_Correlation %>%
  ggplot() + 
  aes(x = naive_VAS,
      y = naive_OT) +
  #Plotting all the raw VAS ratings and OT scores for every listener. 
  #Listeners transcribed and rated 2 out of the 20 speakers, so every listener has 2 pairs of OT/VAS data.
  geom_point(color = my_pal[4], alpha = .7, shape = 1) +
  #Plotting the average OT score and VAS rating per Listener (i.e. the listener's mean OT score with the listener's mean VAS rating)
  geom_point(data = naive_Correlation %>%
               dplyr::group_by(ListenerID) %>%
               dplyr::summarize(naive_VAS = mean(naive_VAS),
                                naive_OT = mean(naive_OT)), 
             color = my_pal[4], size = 2) +
  #Adding the reference line that indicates a theoretically perfect fit. Visualizes if VAS is under- or over-estimating OT scores.
  geom_abline(intercept = 0, slope = 1, linetype = 2) +
  geom_line(data = naive_Correlation %>%
               dplyr::group_by(ListenerID) %>%
               dplyr::summarize(naive_VAS = mean(naive_VAS),
                                naive_OT = mean(naive_OT)),
            method = "lm", stat = "smooth", se = F, color = my_pal[4]) +
  xlab("VAS Ratings") +
  ylab("Orthographic Transcription") +
  coord_cartesian(xlim = c(0,100), ylim = c(0, 100)) +
  theme_classic() +
  theme(aspect.ratio = 1)

```

## Modeling Naïve Listener VAS scores

### Examining SLP EST as a predictor of naive listener VAS.

```{r}

SLP_ESTmodel_VAS <- lm(naive_VAS ~ slp_EST, data = IntSpeakerRatings)
summary(SLP_ESTmodel_VAS)
performance::check_model(SLP_ESTmodel_VAS)

```

### Examining SLP VAS as a predictor of naive listener VAS.

```{r}

SLP_VASmodel_VAS <- lm(naive_VAS ~ slp_VAS, data = IntSpeakerRatings)
summary(SLP_VASmodel_VAS)
performance::check_model(SLP_VASmodel_VAS)

```

