---
title: "SLP Estimations: Data Prep"
output: html_notebook
---

# Packages
```{r}
library(tidyverse) # install.packages('tidyverse')
library(xfun) # install.packages('xfun')
library(remotes) # install.packages('remotes')
library(autoscore) # xfun::install_github("TysonStanley/autoscore")
library(readxl)
library(SnowballC) # install.packages('SnowballC') # this is needed for autoscore
library(corrr)
library(ggpubr)
library(rio)
```

# Naive Listener Data
This section of code uploads, preps, and scores the transcriptions provided by naive listeners.

## Prepping Transcription Data
```{r}
naiveData <- utils::read.csv(paste("Data/Naive Listener Data/",list.files("Data/Naive Listener Data", pattern = "Naive")[1], sep = ""),header = FALSE, stringsAsFactors = FALSE, strip.white=TRUE)
  naiveData[3,] <- NA
  naiveData[2,1:38] <- NA
headings <- as.data.frame(lapply(naiveData[1:3,], paste,collapse="_")) %>%
    str_remove_all("_NA") %>%
    str_remove_all("_4") %>%
    str_remove_all(" ") %>%
    t() %>%
    as.data.frame
    colnames(naiveData) <- headings
    naiveData <- naiveData[-c(1:3),]
    
naiveDataClean <- naiveData %>%
  dplyr::rename(proloficID = Q1.1,
                inUS = Q2.1,
                english = Q2.2,
                famSLHDs = Q2.3,
                age = Q3.1,
                gender = Q3.2,
                race = Q3.3,
                ethnicity = Q3.4) %>%
  dplyr::filter(Finished == "True") %>%
  dplyr::filter(AudioCheck == "Yes, I can hear the audio.") %>%
  dplyr::mutate(ID = paste("N",row_number(),sep = "")) %>% # Creates listener IDs
  relocate(ID, .before = StartDate) # Puts ID col at the beginning

# Removes unwanted variables
rm(headings,naiveData)
```

## Autoscoring the transcriptions
```{r}
PhraseIDs <- readxl::read_xlsx("Data/PhraseID Key.xlsx")
acceptable_spellings <- readxl::read_xlsx("Data/acceptableSpellings.xlsx")

transcriptions <- naiveDataClean %>%
  dplyr::select(ID, AM5_1.1:HDF5_2.VAS) %>%
  tidyr::pivot_longer(!ID,names_to = "file") %>%
  dplyr::rename(Response = value) %>%
  dplyr::filter(!grepl("VAS", file)) %>% # Removes VAS ratings
  dplyr::filter(Response != "") %>% # Removes rows where listeners did not rate
  dplyr::mutate(SpeakerID = sub("_.*", "",file)) %>%
  dplyr::mutate(PhraseID = sub(".*_", "",file)) %>%
  base::merge(PhraseIDs, by = "PhraseID") %>%
  dplyr::mutate(rowID = paste(file,ID, sep = "_")) %>%
  relocate(Response, .after = Target) %>%
  column_to_rownames(var = "rowID")

# Add Target / TargetNumber idiosyncrasies here:
transcriptions$Target[transcriptions$file == "AF1_1.3"] <- "soul has the power of a man half his age"
transcriptions$TargetNumber[transcriptions$file == "AF1_1.3"] <- 10

transcriptions$Target[transcriptions$file == "AF9_2.25"] <- "one hand lay on the lid of his"
transcriptions$TargetNumber[transcriptions$file == "AF9_2.25"] <- 8

transcriptions$Target[transcriptions$file == "HDF5_1.3"] <- "his soul has the power a man half his age"
transcriptions$TargetNumber[transcriptions$file == "HDF5_1.3"] <- 10

transcriptions$Target[transcriptions$file == "HDF5_2.35"] <- "sis, don't get irate, I'll have him home by 8"
transcriptions$TargetNumber[transcriptions$file == "HDF5_2.35"] <- 10

transcriptions$Target[transcriptions$file == "ALSF7_1.12"] <- "but grandma shouts shouts Bereave me"
transcriptions$TargetNumber[transcriptions$file == "ALSF7_1.12"] <- 6

transcriptions$Target[transcriptions$file == "HDF1_1.4"] <- "The way he speaks to my grandma"
transcriptions$TargetNumber[transcriptions$file == "HDF1_1.4"] <- 7

transcriptions$Target[transcriptions$file == "HDF1_1.8"] <- "will learn that she is a ray of light"
transcriptions$TargetNumber[transcriptions$file == "HDF1_1.8"] <- 9

transcriptions$Target[transcriptions$file == "HDF1_1.15"] <- "Grandpa would tease he'd tuck six cakes"
transcriptions$TargetNumber[transcriptions$file == "HDF1_1.15"] <- 7

transcriptions$Target[transcriptions$file == "HDF1_1.16"] <- "in the side of his white coat"
transcriptions$TargetNumber[transcriptions$file == "HDF1_1.16"] <- 7

transcriptions$Target[transcriptions$file == "HDF1_2.2"] <- "It's quite a sight"
transcriptions$TargetNumber[transcriptions$file == "HDF1_2.22"] <- 4

transcriptions$Target[transcriptions$file == "HDF1_2.29"] <- "when he sees his friend"
transcriptions$TargetNumber[transcriptions$file == "HDF1_2.29"] <- 5

transcriptions$Target[transcriptions$file == "HDF1_2.31"] <- "Hey, who let this bore out of the jail"
transcriptions$TargetNumber[transcriptions$file == "HDF1_2.31"] <- 9

transcriptions$Target[transcriptions$file == "HDF1_2.35"] <- "Jeez, don't get irate, I'll have him home by 8"
transcriptions$TargetNumber[transcriptions$file == "HDF1_2.31"] <- 10

transcriptions$Target[transcriptions$file == "PDM12_1.4"] <- "The way he speaks to my grandma"
transcriptions$TargetNumber[transcriptions$file == "PDM12_1.4"] <- 7

transcriptions$Target[transcriptions$file == "PDM12_1.8"] <- "will learn that she is a ray of light"
transcriptions$TargetNumber[transcriptions$file == "PDM12_1.8"] <- 9

transcriptions$Target[transcriptions$file == "PDM12_1.14"] <- "he'd miss the most"
transcriptions$TargetNumber[transcriptions$file == "PDM12_1.14"] <- 4

transcriptions$Target[transcriptions$file == "PDM12_1.15"] <- "Grandpa would often tease he'd duck tuck six cakes"
transcriptions$TargetNumber[transcriptions$file == "PDM12_1.15"] <- 9

transcriptions$Target[transcriptions$file == "PDM12_1.17"] <- "or in my grandma's vase"
transcriptions$TargetNumber[transcriptions$file == "PDM12_1.17"] <- 5

transcriptions$Target[transcriptions$file == "PDM12_2.21"] <- "and give a wailing speech to the kids down at the mall"
transcriptions$TargetNumber[transcriptions$file == "PDM12_2.21"] <- 12

transcriptions$Target[transcriptions$file == "PDM12_2.33"] <- "I'll trust you'll be back later"
transcriptions$TargetNumber[transcriptions$file == "PDM12_2.33"] <- 6

transcriptions$Target[transcriptions$file == "PDM12_2.35"] <- "Jeez, sis, don't get irate, I'll have him back by 8"
transcriptions$TargetNumber[transcriptions$file == "PDM12_2.35"] <- 11


cleanTranscriptions <- transcriptions %>%
  dplyr::select(ID,Target,Response) %>%
  dplyr::rename(id = ID,
                target = Target,
                response = Response)

autoscoreOutput <- autoscore::autoscore(cleanTranscriptions,
                                             acceptable_df = acceptable_spellings,
                                             plural_rule = TRUE,
                                             plural_add_rule = TRUE,
                                             tense_rule = TRUE,
                                             tense_add_rule = TRUE,
                                             a_the_rule = TRUE,
                                             root_word_rule = TRUE,
                                             double_letter_rule = TRUE,
                                             suffix_rule = TRUE)
autoscoreOutput <- autoscoreOutput %>%
  dplyr::rename(Autoscore = autoscore) %>% # This line prepares the autoscored data to be merged with 'transcriptions'
  dplyr::select(Autoscore)
# The below pipe creates the 'scoredTranscriptions', and reorders the columns
 scoredTranscriptions <-  base::merge(autoscoreOutput, transcriptions, by='row.names') %>%
   dplyr::rename(ListenerID = ID) %>%
   relocate(Autoscore, .after = TargetNumber) %>%
   relocate(PhraseID, .after = SpeakerID) %>%
   dplyr::select(!c(Row.names, file))

# Removes unwanted variables
rm(PhraseIDs,acceptable_spellings,transcriptions,cleanTranscriptions,autoscoreOutput)
```

## Transcription Accuracy
```{r}
transAcc <- scoredTranscriptions %>%
  group_by(SpeakerID, ListenerID) %>%
  summarise(sumTarget = sum(TargetNumber),
            sumTranscription = sum(Autoscore),
            .groups = "rowwise") %>%
  dplyr::mutate(transAcc = (sumTranscription/sumTarget)*100) %>%
  dplyr::select(!c(sumTarget,sumTranscription))

Listener_transAcc <- transAcc %>%
  dplyr::group_by(ListenerID) %>%
  dplyr::summarise(transAcc = mean(transAcc))

Speaker_transAcc<- transAcc %>%
  dplyr::group_by(SpeakerID) %>%
  dplyr::summarise(transAcc = mean(transAcc))
```

## VAS 
```{r}
VAS <- naiveDataClean %>%
  dplyr::select(ID, AM5_1.1:HDF5_2.VAS) %>%
  tidyr::pivot_longer(!ID,names_to = "file") %>%
  dplyr::rename(Response = value) %>%
  dplyr::filter(grepl("VAS", file)) %>% # Removes VAS ratings
  dplyr::filter(Response != "") %>% # Removes rows where listeners did not rate
  dplyr::mutate(SpeakerID = sub("_.*", "",file)) %>%
  dplyr::rename(VAS = Response,
                ListenerID = ID) %>%
  relocate(SpeakerID, .before = "ListenerID") %>%
  dplyr::select(!file)

Listener_VAS <- VAS %>%
  dplyr::select(ListenerID, VAS) %>%
  dplyr::mutate(VAS = as.numeric(VAS)) %>%
  dplyr::group_by(ListenerID) %>%
  dplyr::summarise(VAS = mean(VAS))

Speaker_VAS <- VAS %>%
  dplyr::select(SpeakerID, VAS) %>%
  dplyr::mutate(VAS = as.numeric(VAS)) %>%
  dplyr::group_by(SpeakerID) %>%
  dplyr::summarise(VAS = mean(VAS))
```

## Preparing Naive Listener Files for Export
This section preps the NaiveListener_Demographics dataframe, as well a the NaiveListener_correlation dataframe. It also begins the intRatings_bySpeaker dataframe, which will be used for Question #1.

```{r}
# NaiveListener_Demographics ----
ListenerDemo <- naiveDataClean %>%
  dplyr::select(!StartDate:UserLanguage) %>%
  dplyr::select(!proloficID) %>%
  dplyr::select(ID:ethnicity) %>%
  dplyr::rename(ListenerID = ID)

ListenerDemo$race[ListenerDemo$Q3.3_7_TEXT == "Native American/ African amercing"] <- "Biracial or Multiracial"

ListenerDemo <- ListenerDemo %>%
  dplyr::select(!Q3.3_7_TEXT) %>%
  dplyr::select(!Q3.2_8_TEXT)


# NaiveListener_correlation ----
ListenerCorr <- base::merge(transAcc,VAS) %>%
    dplyr::rename(naive_VAS = VAS,
                naive_OT = transAcc)

# intRatings_bySpeaker ----
intRatings <- base::merge(Speaker_VAS,Speaker_transAcc) %>%
  dplyr::rename(naive_VAS = VAS,
                naive_OT = transAcc)

```

## Exporting Files

```{r}
# NaiveListener_Demographics ----
rio::export(ListenerDemo, file = "Data/Shared Data/NaiveListener_Demographics.csv")

# NaiveListener_correlation ----
rio::export(ListenerCorr, file = "Data/Shared Data/NaiveListener_Correlation.csv")

# Removeing unwanted variables
rm(naiveDataClean,Listener_transAcc,Listener_VAS,scoredTranscriptions,Speaker_transAcc,Speaker_VAS,transAcc,VAS)
```

# SLP Data

## Data Cleaning & Preparing Data for Export
This chunk uploads the two SLP dataframes into R, reassigns variable names, and generated slp IDs.

```{r}
# The first SLP survey
slpData1 <- utils::read.csv(paste("Data/SLP Data/",list.files("Data/SLP Data", pattern = "New SLP")[1], sep = ""),header = FALSE, stringsAsFactors = FALSE, strip.white=TRUE)
  slpData1[2,1:39] <- NA
  slpData1[2,281:291] <- NA
  slpData1[3,1:291] <- NA
  slpData1[3,40:79] <- "A"
  slpData1[3,80:119] <- "B"
  slpData1[3,120:159] <- "C"
  slpData1[3,161:200] <- "A"
  slpData1[3,201:240] <- "B"
  slpData1[3,241:280] <- "C"
  headings <- as.data.frame(lapply(slpData1[1:3,], paste,collapse="_")) %>%
    str_remove_all("_NA") %>%
    str_remove_all("_4") %>%
    str_remove_all(" ") %>%
    t() %>%
    as.data.frame
    colnames(slpData1) <- headings
slpData1 <- slpData1[-c(1:3),]
ratings <- slpData1 %>%
  dplyr::select(VAS.AF1_s1_A:EST.PDM9.R_s3_C) %>%
  dplyr::mutate_all(as.numeric)
slpData1[,40:280] <- ratings

# The second SLP survey
slpData2 <- utils::read.csv(paste("Data/SLP Data/",list.files("Data/SLP Data", pattern = "Second SLP")[1], sep = ""),header = FALSE)
    colnames(slpData2) <- headings
slpData2 <- slpData2[-c(1:3),]
ratings <- slpData2 %>%
  dplyr::select(VAS.AF1_s1_A:EST.PDM9.R_s3_C) %>%
  dplyr::mutate_all(as.numeric)
slpData2[,40:280] <- ratings

# Merging the two SLP surveys
slpData <- rbind(slpData1,slpData2)
rm(slpData1, slpData2, ratings, headings) #removing the unneeded two separate SLP surveys.

slpDataClean <- slpData %>%
  dplyr::filter(Finished == "True") %>%
  dplyr::filter(EfficiencyRatings_1 != "") %>% # Filters out the incomplete cases
  dplyr::filter(Email != "") %>%
  dplyr::filter(!str_detect(StartDate, "2021-07-28")) %>% # Filters out the data collected on 7/28, when we got spam bombed
  dplyr::mutate(slpID = paste("slp",row_number(),sep = "")) %>% # Creates SLP IDs
  dplyr::rename(slhDisorder = Q3.1, # Renaming many of the important variables
                citizen = Q3.2,
                english = Q3.3,
                cccs = Q3.4,
                age = Q4.1,
                gender = Q4.2,
                race = Q4.3,
                otherRaceText = Q4.3_7_TEXT,
                ethnicity = Q4.4,
                yearsExp = Q5.1,
                workSetting = Q5.2,
                medSetting = Q5.3,
                ptPopAge = Q5.4,
                percentCaseload = Q5.5,
                perceivedComp = Q5.6_1,
                eff.percentEst = EfficiencyRatings_1,
                eff.vas = EfficiencyRatings_2,
                eff.formal = EfficiencyRatings_3,
                eff.transcription = EfficiencyRatings,
                eff.otherInformal = EfficiencyRatings_5,
                formalAssessments = FormalAssessments,
                otherFormalText = FormalAssessments_9_TEXT,
                otherInformalText = Other) %>%
  relocate(slpID, .before = StartDate) # Puts ID col at the beginning

# These two lines of code fix individual responses to be clean.
  slpDataClean$race[slpDataClean$otherRaceText == "Asian "] <- "Asian American" # This line reclassifies 1 person who identified
  slpDataClean$workSetting[slpDataClean$ResponseId == "R_1JPKUAAUEy5pSF4"] <- "Medical" # This line reclassifies 1 SNF SLP from "Other" to "Medical"
```

## Filtering for Medical SLPs
This is filtering out non-medical SLPs. This is only included for the ASHA poster. Delete this section for manuscript.
```{r}
slpDataClean <- slpDataClean %>%
  dplyr::filter(workSetting == "Medical")
```

## Check for Unique IPs
```{r}
# Checks for unique IPs
if(base::any(duplicated(slpDataClean$IPAddress) == TRUE) == FALSE){
  print("There were no duplicate IP addresses")
} else {
  print("WARNING: Duplicate IP addresses detected")
}
```

## Preparing Data for Export
```{r}
# SLP Demographic Info ----
slpDemo <- slpDataClean %>%
  dplyr::select(!StartDate:Q1.1) %>%
  dplyr::select(!otherRaceText) %>%
  dplyr::select(!Q4.2_8_TEXT) %>%
  dplyr::select(!Q6.1:Q132) %>%
  dplyr::select(!Email:Q125)  

# Prepping the slp_Correlation dataframe ----
slpWide <- slpDataClean %>%
  dplyr::select(slpID, VAS.AF1_s1_A:EST.PDM9.R_s3_C) # This subjects just the ID and the ratings

slpLong <- tidyr::pivot_longer(slpWide, 
                      cols = VAS.AF1_s1_A:EST.PDM9.R_s3_C) # Pivots data to longer
slpLong <- slpLong %>%
  dplyr::rename(fileName = name,
                rating = value)

# Separating the listener's first and second ratings
slpLong_firstRating <- slpLong %>%
  dplyr::filter(!grepl(".R",fileName))

slpLong_secondRating <- slpLong %>%
  dplyr::filter(grepl(".R",fileName)) %>%
  dplyr::rename(relRating = rating) %>%
  dplyr::mutate(newName = str_replace(fileName,".R","")) %>%
  dplyr::select(!fileName) %>%
  dplyr::rename(fileName = newName)

# Merging the listeners first and second ratings
slpLong <- base::merge(slpLong_firstRating,
                    slpLong_secondRating,
                    by = intersect(names(slpLong_firstRating), names(slpLong_secondRating)),
                    no.dups = FALSE) %>%
  dplyr::mutate(Speaker = sub("_.*", "", fileName),
                Speaker = sub("VAS.", "", Speaker),
                Speaker = sub("EST.", "", Speaker),
                Method = substr(fileName,1,3)) %>%
  dplyr::filter(rating != "") %>%
  dplyr::select(!fileName) %>%
  dplyr::relocate(Speaker, .after = slpID) %>%
  dplyr::relocate(Method, .after = Speaker)

slpLong_VAS <- slpLong %>%
  dplyr::filter(Method == "VAS") %>%
  dplyr::select(!Method) %>%
  dplyr::rename(slp_VAS = rating,
                slp_VAS.Rel = relRating)

slpLong_EST <- slpLong %>%
  dplyr::filter(Method == "EST") %>%
  dplyr::select(!Method) %>%
  dplyr::rename(slp_EST = rating,
                slp_EST.Rel = relRating)

slpCorr <- base::merge(slpLong_VAS,slpLong_EST, by = intersect(names(slpLong_VAS),names(slpLong_EST)),all = TRUE)

# Prepping the SLP portion of intRatings dataframe ----
slpRatings_bySpeaker <- slpLong %>%
  dplyr::select(!c(slpID, relRating)) %>%
  dplyr::group_by(Speaker, Method) %>%
  dplyr::summarise(rating = mean(rating),
                   .groups = "rowwise")

slpRatings_bySpeaker <- pivot_wider(slpRatings_bySpeaker, names_from = Method, values_from = rating) %>%
  dplyr::rename(slp_EST = EST,
                slp_VAS = VAS,
                SpeakerID = Speaker)

intRatings <- base::merge(intRatings,slpRatings_bySpeaker)

```

## Exporting Files
```{r}
# SLP Demographics
rio::export(slpDemo, file = "Data/Shared Data/SLP_Demographics.csv")

# SLP Correlation
rio::export(slpCorr, file = "Data/Shared Data/SLP_Correlation.csv")

# SpeakerIntelligibility
rio::export(intRatings, file="Data/Shared Data/SpeakerIntelligibility.csv")

# Removing unwanted variables
rm(slpData, slpDataClean,slpLong,slpLong_EST,slpLong_firstRating,slpLong_secondRating,slpLong_VAS,slpRatings_bySpeaker,slpWide)
```


